---
title: "Test Cases for Programming Assignment 2"
author: "Stephen Schwartz"
date: "Sunday, October 26, 2014"
output: word_document
---

Following are the declarations for the two functions in 'cachematrx.R' and then 3 test cases that demonstrate the functions.


Function Calls:
```{r}
## this function takes an input matrix, x, and then stores it and it's inverse 
## in two global values for quick calculations later.

makeCacheMatrix <- function(x) {
  
  ##  1. store a matrix
  ##  2. store the inverse of the current matrix
    
  ## store the current matrix for comparison in higher function
  ## using <<- so that matrixStore can be referred to globally
  matrixStore <<- x
  ## calculate the inverse for the current matrix
  matrixInverseStore <<- solve(matrixStore)  
}


## This function takes an input matrix, x, and then checks to see if:
## 1. Is this matrix invertible?  If not, will spit out an error msg.
## 2. If it is invertible, then it will check to see if it's been previously processed 
##    and exists in the cache.  If it does, values are pulled from cache.
## 3. if not, then makeCacheMatrix is called to generate the inverse.

cacheSolve <- function(x) {
  ### x is a user defined matrix
  ### returns an inverse of the matrix x if it's invertible, otherwise and error msg

  ## initialize matrixStore if it doesn't exist
  ## need to figure out nrows and ncols from input matrix to initialize if necessary
  if (!exists("matrixStore")) {matrixStore <- matrix(c(1),nrow=nrow(x),ncol=ncol(x))}
  
  ### check to see if it's invertible
  if (round(det(x),0) == 0) {
    message("Matrix isn't invertible.  Please try again.")
  } else{ 
    ### if it's invertable then continue
         
         message("Matrix is invertible.  Processing continues..")
          ### check to see if it's already been declared
         message("Checking...")
         
          ## check to see if it matches cached matrix
         if (!all(matrixStore == x)){
              makeCacheMatrix(x)
      
              message("First time this matrix was inverted.  Inverse calculated and cached values set.")
              matrixInverse <- matrixInverseStore
              return(matrixInverse)
          } else{
              message("This matrix has been previously processed, retrieving from cache...")
              matrixInverse <- matrixInverseStore
              return(matrixInverse)
          }
  }
}
```

**Below are the test cases:**

*Test case 1* generates a simple matrix 3x3 that IS invertible and then calls the function 2x to demonstrate cache extraction.
```{r}
myMatrix <- matrix(c(-1,2,3,0,1,0,-1,0,1), 
            nrow=3,
            ncol=3,
            byrow=TRUE)
dimnames(myMatrix) = list(
  c("row1","row2","row3"),
  c("col1","col2","col3"))

## first time called, should generate
cacheSolve(myMatrix)

## second time called, should pull from cache
cacheSolve(myMatrix)
```


*Test case 2* generates with with a slight modification to myMatrix to see if it re-generates the inverse rather than returning the cache.  Note, this matrix IS invertible.

```{r}
myMatrix <- matrix(c(-1,2,4,0,1,0,-1,0,1), 
                   nrow=3,
                   ncol=3,
                   byrow=TRUE)
dimnames(myMatrix) = list(
  c("row1","row2","row3"),
  c("col1","col2","col3"))

## first time called, should generate
cacheSolve(myMatrix)

## second time called, should pull from cache
cacheSolve(myMatrix)
```

*Test case 3* generatse a simple matrix 3x3 that is not invertible, which should result in an error msg.
```{r}
myMatrix <- matrix(c(1,2,3,4,5,6,7,8,9), 
            nrow=3,
            ncol=3,
            byrow=TRUE)
dimnames(myMatrix) = list(
  c("row1","row2","row3"),
  c("col1","col2","col3"))

## call this one, should kick out an error msg
cacheSolve(myMatrix)
```